---
title: "Statistical Learning"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(glmnet)
set.seed(11)
```

##Lasso

To illustrate the lasso, we’ll data from a study of factors that affect birthweight. The code chunk below loads and cleans these data, converts to factors where appropriate, and takes a sample of size 200 from the result.

```{r}
bwt_df = 
  read_csv("data/birthweight.csv") |> 
  janitor::clean_names() |>
  mutate(
    babysex = as.factor(babysex),
    babysex = fct_recode(babysex, "male" = "1", "female" = "2"),
    frace = as.factor(frace),
    frace = fct_recode(
      frace, "white" = "1", "black" = "2", "asian" = "3", 
      "puerto rican" = "4", "other" = "8"),
    malform = as.logical(malform),
    mrace = as.factor(mrace),
    mrace = fct_recode(
      mrace, "white" = "1", "black" = "2", "asian" = "3", 
      "puerto rican" = "4")) |> 
  sample_n(200)

bwt_df
```

get predictors and outcome.

To fit a lasso model, we’ll use glmnet. This package is widely used and broadly useful, but predates the tidyverse by a long time. The interface asks for an outcome vector y and a matrix of predictors X, which are created next. To create a predictor matrix that includes relevant dummy variables based on factors, we’re using model.matrix and excluding the intercept

```{r}
x = model.matrix(bwt ~ ., bwt_df)[,-1]
y = bwt_df |> pull(bwt)
```

We fit the lasso model for each tuning parameter in a pre-defined grid lambda, and then compare the fits using cross validation. I chose this grid using the trusty “try things until it looks right” method; glmnet will pick something reasonable by default if you prefer that.

```{r}
lambda = 10^(seq(3, -2, -0.1))

lasso_fit =
  glmnet(x, y, lambda = lambda)

lasso_cv =
  cv.glmnet(x, y, lambda = lambda)

lambda_opt = lasso_cv$lambda.min
```
lasso_cv = we are doing lasso cross validation 

The plot below shows coefficient estimates corresponding to a subset of the predictors in the dataset – these are predictors that have non-zero coefficients for at least one tuning parameter value in the pre-defined grid. As lambda increases, the coefficient values are shrunk to zero and the model becomes more sparse. The optimal tuning parameter, determined using cross validation, is shown by a vertical blue line.

```{r}
lasso_fit |> 
  broom::tidy() |> 
  select(term, lambda, estimate) |> 
  complete(term, lambda, fill = list(estimate = 0) ) |> 
  filter(term != "(Intercept)") |> 
  ggplot(aes(x = log(lambda, 10), y = estimate, group = term, color = term)) + 
  geom_path() + 
  geom_vline(xintercept = log(lambda_opt, 10), color = "blue", size = 1.2) +
  theme(legend.position = "none")
```

show the CV result 
This is relatively shallow – having nothing at all in your model isn’t great, but you can get reasonable predictions from models that have “too many” predictors.

```{r}
lasso_cv |> 
  broom::tidy() |> 
  ggplot(aes(x = log(lambda, 10), y = estimate)) + 
  geom_point()  
```





